---
title: "Project2_decision_tree"
author: "Reha"
date: "02/05/2021"
output: html_document
---

```{r setup, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
knitr::opts_chunk$set(warning = F, results = 'markup', message = F)
# knitr::opts_chunk$set(warning = F, results = 'hide', message = F)
# knitr::opts_chunk$set(include = F)
# knitr::opts_chunk$set(echo = TRUE)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

```{r basic, include=F}
# use this function to conveniently load libraries and work smoothly with knitting
# can add quietly=T option to the require() function
loadPkg = function(pkg, character.only = FALSE) { 
  if (!character.only) { pkg <- as.character(substitute(pkg)) }
  pkg <- ifelse(!character.only, as.character(substitute(pkg)) , pkg)  
  if (!require(pkg,character.only=T, quietly =T)) {  install.packages(substitute(pkg),dep=T); if(!require(pkg,character.only=T)) stop("Package not found") } 
}
loadPkg(knitr)
# unload/detact package when done using it
unloadPkg = function(pkg, character.only = FALSE) { 
  if(!character.only) { pkg <- as.character(substitute(pkg)) } 
  search_item <- paste("package", pkg,sep = ":") 
  while(search_item %in% search()) { detach(search_item, unload = TRUE, character.only = TRUE) } 
}
loadPkg("xtable")
loadPkg("kableExtra")
loadPkg("stringi")
xkabledply = function(modelsmmrytable, title="Table", digits = 4, pos="left", bso="striped", wide=FALSE) { 
  #' Combining base::summary, xtable, and kableExtra, to easily display model summary. 
  #' wrapper for the base::summary function on model objects
  #' Can also use as head for better display
  #' ELo 202004 GWU DATS
  #' version 1.2
  #' @param modelsmmrytable This can be a generic table, a model object such as lm(), or the summary of a model object summary(lm()) 
  #' @param title Title of table. 
  #' @param digits Number of digits to display
  #' @param pos Position of table, c("left","center","right") 
  #' @param bso bootstrap_options = c("basic", "striped", "bordered", "hover", "condensed", "responsive")
  #' @param wide print table in long (FALSE) format or wide (TRUE) format
  #' @return HTML table for display
  #' @examples
  #' library("xtable")
  #' library("kableExtra")
  #' xkabledply( df, title="Table testing", pos="left", bso="hover" )
  #' xkabledply( ISLR::Hitters[1:5,] )
  if (wide) { modelsmmrytable <- t(modelsmmrytable) }
  modelsmmrytable %>%
    xtable() %>% 
    kable(caption = title, digits = digits) %>%
    kable_styling(bootstrap_options = bso, full_width = FALSE, position = pos)
}
xkabledplyhead = function(df, rows=5, title="Head", digits = 4, pos="left", bso="striped") { 
  xkabledply(df[1:rows, ], title, digits, pos, bso, wide=FALSE)
}
xkabledplytail = function(df, rows=5, title="Tail", digits = 4, pos="left", bso="striped") { 
  trows = nrow(df)
  xkabledply(df[ (trows-rows+1) : trows, ], title, digits, pos, bso, wide=FALSE)
}
xkablesummary = function(df, title="Table: Statistics summary.", digits = 4, pos="left", bso="striped") { 
  #' Combining base::summary, xtable, and kableExtra, to easily display numeric variable summary of dataframes. 
  #' ELo 202004 GWU DATS
  #' version 1.2
  #' @param df The dataframe.
  #' @param title Title of table. 
  #' @param digits Number of digits to display
  #' @param pos Position of table, c("left","center","right") 
  #' @param bso bootstrap_options = c("basic", "striped", "bordered", "hover", "condensed", "responsive")
  #' @return The HTML summary table for display, or for knitr to process into other formats 
  #' @examples
  #' xkablesummary( faraway::ozone )
  #' xkablesummary( ISLR::Hitters, title="Five number summary", pos="left", bso="hover"  )
  
  s = summary(df) %>%
    apply( 2, function(x) stringr::str_remove_all(x,c("Min.\\s*:\\s*","1st Qu.\\s*:\\s*","Median\\s*:\\s*","Mean\\s*:\\s*","3rd Qu.\\s*:\\s*","Max.\\s*:\\s*")) ) %>% # replace all leading words
    apply( 2, function(x) stringr::str_trim(x, "right")) # trim trailing spaces left
  
  colnames(s) <- stringr::str_trim(colnames(s))
  
  if ( dim(s)[1] ==6 ) { rownames(s) <- c('Min','Q1','Median','Mean','Q3','Max') 
  } else if ( dim(s)[1] ==7 ) { rownames(s) <- c('Min','Q1','Median','Mean','Q3','Max','NA') }
  
  xkabledply(s, title=title, digits = digits, pos=pos, bso=bso )
}
xkablevif = function(model, title="VIFs of the model", digits = 3, pos="left", bso="striped", wide=TRUE) { 
  #' Combining faraway::vif, xtable, and kableExtra, to easily display numeric summary of VIFs for a model. 
  #' ELo 202004 GWU DATS
  #' version 1.2
  #' @param model The lm or compatible model object.
  #' @param title Title of table. 
  #' @param digits Number of digits to display
  #' @param pos Position of table, c("left","center","right") 
  #' @param bso bootstrap_options = c("basic", "striped", "bordered", "hover", "condensed", "responsive")
  #' @param wide print table in long (FALSE) format or wide (TRUE) format
  #' @return The HTML summary table of the VIFs for a model for display, or for knitr to process into other formats 
  #' @examples
  #' xkablevif( lm(Salary~Hits+RBI, data=ISLR::Hitters), wide=T )
  
  vifs = table( names(model$coefficients)[2:length(model$coefficients)] ) # remove intercept to set column names
  vifs[] = faraway::vif(model) # set the values
  if (wide) { vifs <- t(vifs) }
  xkabledply( vifs, title=title, digits = digits, pos=pos, bso=bso )
}
```


```{r}
data_raw <-data.frame(read.csv("telecom_users (1).csv"))
summary(data_raw)
head(data_raw)
```


# Preliminary Data Cleaning

How many N/A are there?
```{r}
data_raw[!complete.cases(data_raw),]
```

10 Missing Values in `TotalCharges` column. We can drop them.
```{r}
telco_data <- na.omit(data_raw)
nrow(telco_data)
```

We don't need customer ID. Drop first two columns
```{r}
telco_data <- telco_data[3:22]
head(teleco_data)
```


Let's convert variables into factors
```{r}
telco_data$gender <- as.factor(telco_data$gender)
telco_data$SeniorCitizen <- as.factor(telco_data$SeniorCitizen)
telco_data$Partner <- as.factor(telco_data$Partner)
telco_data$Dependents <- as.factor(telco_data$Dependents)
telco_data$PhoneService <- as.factor(telco_data$PhoneService)
telco_data$MultipleLines <- as.factor(telco_data$MultipleLines)
telco_data$InternetService <- as.factor(telco_data$InternetService)
telco_data$OnlineSecurity <- as.factor(telco_data$OnlineSecurity)
telco_data$OnlineBackup <- as.factor(telco_data$OnlineBackup)
telco_data$DeviceProtection <- as.factor(telco_data$DeviceProtection)
telco_data$TechSupport <- as.factor(telco_data$TechSupport)
telco_data$StreamingTV <- as.factor(telco_data$StreamingTV)
telco_data$StreamingMovies <- as.factor(telco_data$StreamingMovies)
telco_data$Contract <- as.factor(telco_data$Contract)
telco_data$PaperlessBilling <- as.factor(telco_data$PaperlessBilling)
telco_data$PaymentMethod <- as.factor(telco_data$PaymentMethod)
telco_data$Churn <- as.factor(telco_data$Churn)
```

Decision Tree:
```{r results='markup'}
set.seed(1)
treefit1<-rpart(Churn ~., data=telco_data, method="class",control = list(maxdepth = 5) )
```
```{r}
printcp(treefit1) # display the results 
plotcp(treefit1) # visualize cross-validation results 
summary(treefit1) # detailed summary of splits
```
```{r}
plot(treefit1, uniform=TRUE, main="Classification Tree for Churn")
text(treefit1, use.n=TRUE, all=TRUE, cex=.8)
```
```{r}
post(treefit1, file = "decisionTree1.ps", title = "Classification Tree for Churn")
```

```{r}
loadPkg("rpart")
loadPkg("rpart.plot")
rpart.plot(treefit1)
```

So here are the results. 
We see that at the first branching point, there are 4389 no and   1587 yes(with churn). The first split yields 2707 outcomes with contract= one or two years and 3269 outcomes with contract= month to month. 

We can also use some handy library to calculate these percentages in the confusion matrix.


```{r}
loadPkg("caret") 
cm = confusionMatrix( predict(treefit1, type = "class"), reference = telco_data[, "Churn"] )
print('Overall: ')
cm$overall
print('Class: ')
cm$byClass
unloadPkg("caret")
```


```{r}
xkabledply(cm$table, "confusion matrix")
```
Let us try different different maxdepths, and collect the result summaries for display.
```{r}
loadPkg("rpart")
loadPkg("caret")


# create an empty dataframe to store the results from confusion matrices
confusionMatrixResultDf = data.frame( Depth=numeric(0), Accuracy= numeric(0), Sensitivity=numeric(0), Specificity=numeric(0), Pos.Pred.Value=numeric(0), Neg.Pred.Value=numeric(0), Precision=numeric(0), Recall=numeric(0), F1=numeric(0), Prevalence=numeric(0), Detection.Rate=numeric(0), Detection.Prevalence=numeric(0), Balanced.Accuracy=numeric(0), row.names = NULL )

for (deep in 2:10) {
  treefit2 <- rpart(Churn ~., data=telco_data, method="class", control = list(maxdepth = deep) )
  # 
  cm = confusionMatrix( predict(treefit2, type = "class"), reference = telco_data[, "Churn"] ) # from caret library
  # 
  cmaccu = cm$overall['Accuracy']
  # print( paste("Total Accuracy = ", cmaccu ) )
  # 
  cmt = data.frame(Depth=deep, Accuracy = cmaccu, row.names = NULL ) # initialize a row of the metrics 
  cmt = cbind( cmt, data.frame( t(cm$byClass) ) ) # the dataframe of the transpose, with k valued added in front
  confusionMatrixResultDf = rbind(confusionMatrixResultDf, cmt)
  # print("Other metrics : ")
}

unloadPkg("caret")
```
```{r}
xkabledply(confusionMatrixResultDf, title="churn Classification Trees summary with varying MaxDepth")
```

The model's accuracy is not improving even when the depth is increased.

```{r}
loadPkg("rattle")
fancyRpartPlot(treefit1)
```


Let's prune the model to understand the optimal decision tree for the dataset.
```{r}
#prune the tree 
ptree <- prune(treefit1, cp = treefit1$cptable[2,"CP"])


# plot the pruned tree 
fancyRpartPlot(ptree)
# For boring plot, use codes below instead
plot(ptree, uniform=TRUE, main="Pruned Classification Tree for Churn")
text(ptree, use.n=TRUE, all=TRUE, cex=.8)
```

