---
title: "Telecom User Churn Prediction"
author: "KONG Deyu"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    # number_sections: true
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
knitr::opts_chunk$set(warning = F, results = 'markup', message = F)
# knitr::opts_chunk$set(warning = F, results = 'hide', message = F)
# knitr::opts_chunk$set(include = F)
# knitr::opts_chunk$set(echo = TRUE)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

```{r basic, include=F}
# use this function to conveniently load libraries and work smoothly with knitting
# can add quietly=T option to the require() function
loadPkg = function(pkg, character.only = FALSE) { 
  if (!character.only) { pkg <- as.character(substitute(pkg)) }
  pkg <- ifelse(!character.only, as.character(substitute(pkg)) , pkg)  
  if (!require(pkg,character.only=T, quietly =T)) {  install.packages(substitute(pkg),dep=T); if(!require(pkg,character.only=T)) stop("Package not found") } 
}
loadPkg(knitr)

# unload/detact package when done using it
unloadPkg = function(pkg, character.only = FALSE) { 
  if(!character.only) { pkg <- as.character(substitute(pkg)) } 
  search_item <- paste("package", pkg,sep = ":") 
  while(search_item %in% search()) { detach(search_item, unload = TRUE, character.only = TRUE) } 
}

loadPkg("xtable")
loadPkg("kableExtra")
loadPkg("stringi")

xkabledply = function(modelsmmrytable, title="Table", digits = 4, pos="left", bso="striped", wide=FALSE) { 
  #' Combining base::summary, xtable, and kableExtra, to easily display model summary. 
  #' wrapper for the base::summary function on model objects
  #' Can also use as head for better display
  #' ELo 202004 GWU DATS
  #' version 1.2
  #' @param modelsmmrytable This can be a generic table, a model object such as lm(), or the summary of a model object summary(lm()) 
  #' @param title Title of table. 
  #' @param digits Number of digits to display
  #' @param pos Position of table, c("left","center","right") 
  #' @param bso bootstrap_options = c("basic", "striped", "bordered", "hover", "condensed", "responsive")
  #' @param wide print table in long (FALSE) format or wide (TRUE) format
  #' @return HTML table for display
  #' @examples
  #' library("xtable")
  #' library("kableExtra")
  #' xkabledply( df, title="Table testing", pos="left", bso="hover" )
  #' xkabledply( ISLR::Hitters[1:5,] )
  if (wide) { modelsmmrytable <- t(modelsmmrytable) }
  modelsmmrytable %>%
    xtable() %>% 
    kable(caption = title, digits = digits) %>%
    kable_styling(bootstrap_options = bso, full_width = FALSE, position = pos)
}

xkabledplyhead = function(df, rows=5, title="Head", digits = 4, pos="left", bso="striped") { 
  xkabledply(df[1:rows, ], title, digits, pos, bso, wide=FALSE)
}

xkabledplytail = function(df, rows=5, title="Tail", digits = 4, pos="left", bso="striped") { 
  trows = nrow(df)
  xkabledply(df[ (trows-rows+1) : trows, ], title, digits, pos, bso, wide=FALSE)
}

xkablesummary = function(df, title="Table: Statistics summary.", digits = 4, pos="left", bso="striped") { 
  #' Combining base::summary, xtable, and kableExtra, to easily display numeric variable summary of dataframes. 
  #' ELo 202004 GWU DATS
  #' version 1.2
  #' @param df The dataframe.
  #' @param title Title of table. 
  #' @param digits Number of digits to display
  #' @param pos Position of table, c("left","center","right") 
  #' @param bso bootstrap_options = c("basic", "striped", "bordered", "hover", "condensed", "responsive")
  #' @return The HTML summary table for display, or for knitr to process into other formats 
  #' @examples
  #' xkablesummary( faraway::ozone )
  #' xkablesummary( ISLR::Hitters, title="Five number summary", pos="left", bso="hover"  )
  
  s = summary(df) %>%
    apply( 2, function(x) stringr::str_remove_all(x,c("Min.\\s*:\\s*","1st Qu.\\s*:\\s*","Median\\s*:\\s*","Mean\\s*:\\s*","3rd Qu.\\s*:\\s*","Max.\\s*:\\s*")) ) %>% # replace all leading words
    apply( 2, function(x) stringr::str_trim(x, "right")) # trim trailing spaces left
  
  colnames(s) <- stringr::str_trim(colnames(s))
  
  if ( dim(s)[1] ==6 ) { rownames(s) <- c('Min','Q1','Median','Mean','Q3','Max') 
  } else if ( dim(s)[1] ==7 ) { rownames(s) <- c('Min','Q1','Median','Mean','Q3','Max','NA') }
  
  xkabledply(s, title=title, digits = digits, pos=pos, bso=bso )
}

xkablevif = function(model, title="VIFs of the model", digits = 3, pos="left", bso="striped", wide=TRUE) { 
  #' Combining faraway::vif, xtable, and kableExtra, to easily display numeric summary of VIFs for a model. 
  #' ELo 202004 GWU DATS
  #' version 1.2
  #' @param model The lm or compatible model object.
  #' @param title Title of table. 
  #' @param digits Number of digits to display
  #' @param pos Position of table, c("left","center","right") 
  #' @param bso bootstrap_options = c("basic", "striped", "bordered", "hover", "condensed", "responsive")
  #' @param wide print table in long (FALSE) format or wide (TRUE) format
  #' @return The HTML summary table of the VIFs for a model for display, or for knitr to process into other formats 
  #' @examples
  #' xkablevif( lm(Salary~Hits+RBI, data=ISLR::Hitters), wide=T )
  
  vifs = table( names(model$coefficients)[2:length(model$coefficients)] ) # remove intercept to set column names
  vifs[] = faraway::vif(model) # set the values
  if (wide) { vifs <- t(vifs) }
  xkabledply( vifs, title=title, digits = digits, pos=pos, bso=bso )
}
```


```{r}
data_raw <- read.csv("telecom_users.csv")
summary(data_raw)
head(data_raw)
```


# Preliminary Data Cleaning

How many N/A are there?
```{r}
data_raw[!complete.cases(data_raw),]
```

10 Missing Values in `TotalCharges` column. We can drop them.
```{r}
df <- na.omit(data_raw)
nrow(df)
```

We don't need customer ID. Drop first two columns
```{r}
df <- df[3:22]
head(df)
```


Encode `SeniorCitizen` status to factor for regression, and `Churn` to binary
```{r}
df$SeniorCitizen <- ifelse(df$SeniorCitizen == 1, "Yes", "No")
df$Churn <- ifelse(df$Churn == "Yes", 1, 0)
```


```{r}
df_Churn <- with(df, subset(df, Churn == 1))
nrow(df_Churn)
df_Stay <- with(df, subset(df, Churn == 0))
nrow(df_Stay)
```

Does `TotalCharges` affect Churn decision?
```{r}
t.test(df_Churn$TotalCharges, df_Stay$TotalCharges)
```

Does `tenure` affect Churn decision?
```{r}
t.test(df_Churn$tenure, df_Stay$tenure)
```

Does the type of `Contract` affect Churn decision?
```{r}
chisq.test(xtabs(~ Churn + Contract, data = df))
```

`gender`?
```{r}
chisq.test(xtabs(~ Churn + gender, data = df))
```

`SeniorCitizen`?
```{r}
chisq.test(xtabs(~ Churn + SeniorCitizen, data = df))
```

`Partner`?
```{r}
chisq.test(xtabs(~ Churn + Partner, data = df))
```

`Dependents`?
```{r}
chisq.test(xtabs(~ Churn + Dependents, data = df))
```



Model Building:
```{r}
glm1 <- glm(Churn ~ TotalCharges + tenure + Contract, data = df, family = "binomial")
summary(glm1)
```

```{r}
coef_glm1 <- exp(coef(glm1))
coef_glm1
```

Obviously, higher the charges, the more likely customer churn. Long term customer have better loyalty.

Model Evaluation: Confusion Matrix
```{r}
loadPkg("caret")
pred1 <- fitted(glm1)
pred1 <- ifelse(pred1 > 0.5, 1, 0) %>% factor()
confusionMatrix(data = pred1, factor(df$Churn))
```

Hosmer and Lemeshow test
```{r}
loadPkg("ResourceSelection")
hoslem.test(df$Churn, fitted(glm1))
```

Obviously, our model is not good enough

```{r}
loadPkg("pROC")
prob1 <- predict(glm1, type = "response")
df$prob <- prob1
h1 <- roc(Churn ~ prob, data = df)
auc(h1)
plot(h1)
```

AUC of 0.819, we can improve from here.

McFadden
```{r}
loadPkg("pscl")
pR2_glm1 <- pR2(glm1)
pR2_glm1[4]
```

Are people more sensitive to marginal(Monthly) expense?
```{r}
glm2 <- glm(Churn ~ MonthlyCharges + tenure + Contract, data = df, family = "binomial")
summary(glm2)
```

```{r}
coef_glm2 <- exp(coef(glm2))
coef_glm2
```


Does Phone and Internet Services affect `Churn` Decision?
```{r}
glm3 <- glm(Churn ~ MonthlyCharges + tenure + Contract + PhoneService + InternetService, data = df, family = "binomial")
summary(glm3)
```


```{r}
coef_glm3 <- exp(coef(glm3))
coef_glm3
```

Fiber internet service must be laggy and expensive. So many customers are leaving after using it.

Again, Model evaluations:
Confusion Matrix
```{r}
pred3 <- fitted(glm3)
pred3 <- ifelse(pred3 > 0.5, 1, 0) %>% factor()
confusionMatrix(data = pred3, factor(df$Churn))
```

Hosmer and Lemeshow test
```{r}
hoslem.test(df$Churn, fitted(glm3))
```

Still not good enough

```{r}
prob3 <- predict(glm3, type = "response" )
df$prob <- prob3
h3 <- roc(Churn ~ prob, data = df)
auc(h3)
plot(h3)
```

Some improvement.

McFadden
```{r}
pR2_glm3 <- pR2(glm3)
pR2_glm3[4]
```


```{r}
glm4 <- glm(Churn ~ MonthlyCharges + tenure + Contract + PhoneService + InternetService + SeniorCitizen + Partner + Dependents, data = df, family = "binomial")
summary(glm4)
```


```{r}
glm5 <- glm(Churn ~ MonthlyCharges + tenure + Contract + PhoneService + MultipleLines + InternetService + OnlineSecurity + OnlineBackup + DeviceProtection + TechSupport + StreamingTV + StreamingMovies + SeniorCitizen, data = df, family = "binomial")
summary(glm5)
```


```{r}
glm6 <- glm(Churn ~ MonthlyCharges + tenure + Contract + PhoneService + MultipleLines + InternetService + PaperlessBilling + PaymentMethod + SeniorCitizen, data = df, family = "binomial")
summary(glm6)
```

```{r}
coef_glm6 <- exp(coef(glm6))
coef_glm6
```

Again, Model evaluations:
Confusion Matrix
```{r}
pred6 <- fitted(glm6)
pred6 <- ifelse(pred6 > 0.5, 1, 0) %>% factor()
confusionMatrix(data = pred6, factor(df$Churn))
```

Hosmer and Lemeshow test
```{r}
hoslem.test(df$Churn, fitted(glm6))
```

p-value big enough

```{r}
prob6 <- predict(glm6, type = "response" )
df$prob <- prob6
h6 <- roc(Churn ~ prob, data = df)
auc(h6)
plot(h6)
```

Some improvement.

McFadden
```{r}
pR2_glm6 <- pR2(glm6)
pR2_glm6[4]
```

Finnaly, compare to full model. Does full model performs better?
```{r}
glmF <- glm(Churn ~ ., data = df, family = "binomial")
summary(glmF)
```

```{r}
predF <- fitted(glmF)
predF <- ifelse(predF > 0.5, 1, 0) %>% factor()
confusionMatrix(data = predF, factor(df$Churn))
```

```{r}
probF <- predict(glmF, type = "response" )
df$prob <- probF
hF <- roc(Churn ~ prob, data = df)
auc(hF)
plot(hF)
```

```{r}
pR2_glmF <- pR2(glmF)
pR2_glmF[4]
```

Summary: Full model improves prediction accuracy and False Negative rates, improves pseudo $R^2$ and `AUC`, but harms interpretation.



