---
title: "NY Child Care_EAD"
author: 
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    # number_sections: true
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r basic, include=F}
# use this function to conveniently load libraries and work smoothly with knitting
# can add quietly=T option to the require() function
# the loadPkg function essentially replaced/substituted two functions install.packages() and library() in one step.
loadPkg = function(x) { if (!require(x,character.only=T, quietly =T)) { install.packages(x,dep=T,repos="http://cran.us.r-project.org"); if(!require(x,character.only=T)) stop("Package not found") } }

# unload/detact package when done using it
unloadPkg = function(pkg, character.only = FALSE) { 
  if(!character.only) { pkg <- as.character(substitute(pkg)) } 
  search_item <- paste("package", pkg,sep = ":") 
  while(search_item %in% search()) { detach(search_item, unload = TRUE, character.only = TRUE) } 
}
```

```{r setup, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
knitr::opts_chunk$set(warning = F, results = "markup", message = F)
# knitr::opts_chunk$set(warning = F, results = "hide", message = F)
# knitr::opts_chunk$set(include = F)
# knitr::opts_chunk$set(echo = TRUE)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

```{r xkablesummary}
loadPkg("xtable")
loadPkg("kableExtra")
loadPkg("stringi")

xkabledply = function(modelsmmrytable, title="Table", digits = 4, pos="left", bso="striped", wide=FALSE) { 
  #' Combining base::summary, xtable, and kableExtra, to easily display model summary. 
  #' wrapper for the base::summary function on model objects
  #' Can also use as head for better display
  #' ELo 202004 GWU DATS
  #' version 1.2
  #' @param modelsmmrytable This can be a generic table, a model object such as lm(), or the summary of a model object summary(lm()) 
  #' @param title Title of table. 
  #' @param digits Number of digits to display
  #' @param pos Position of table, c("left","center","right") 
  #' @param bso bootstrap_options = c("basic", "striped", "bordered", "hover", "condensed", "responsive")
  #' @param wide print table in long (FALSE) format or wide (TRUE) format
  #' @return HTML table for display
  #' @examples
  #' library("xtable")
  #' library("kableExtra")
  #' xkabledply( df, title="Table testing", pos="left", bso="hover" )
  #' xkabledply( ISLR::Hitters[1:5,] )
  if (wide) { modelsmmrytable <- t(modelsmmrytable) }
  modelsmmrytable %>%
    xtable() %>% 
    kable(caption = title, digits = digits) %>%
    kable_styling(bootstrap_options = bso, full_width = FALSE, position = pos)
}

xkabledplyhead = function(df, rows=5, title="Head", digits = 4, pos="left", bso="striped") { 
  xkabledply(df[1:rows, ], title, digits, pos, bso, wide=FALSE)
}

xkabledplytail = function(df, rows=5, title="Tail", digits = 4, pos="left", bso="striped") { 
  trows = nrow(df)
  xkabledply(df[ (trows-rows+1) : trows, ], title, digits, pos, bso, wide=FALSE)
}

xkablesummary = function(df, title="Table: Statistics summary.", digits = 4, pos="left", bso="striped") { 
  #' Combining base::summary, xtable, and kableExtra, to easily display numeric variable summary of dataframes. 
  #' ELo 202004 GWU DATS
  #' version 1.2
  #' @param df The dataframe.
  #' @param title Title of table. 
  #' @param digits Number of digits to display
  #' @param pos Position of table, c("left","center","right") 
  #' @param bso bootstrap_options = c("basic", "striped", "bordered", "hover", "condensed", "responsive")
  #' @return The HTML summary table for display, or for knitr to process into other formats 
  #' @examples
  #' xkablesummary( faraway::ozone )
  #' xkablesummary( ISLR::Hitters, title="Five number summary", pos="left", bso="hover"  )
  
  s = summary(df) %>%
    apply( 2, function(x) stringr::str_remove_all(x,c("Min.\\s*:\\s*","1st Qu.\\s*:\\s*","Median\\s*:\\s*","Mean\\s*:\\s*","3rd Qu.\\s*:\\s*","Max.\\s*:\\s*")) ) %>% # replace all leading words
    apply( 2, function(x) stringr::str_trim(x, "right")) # trim trailing spaces left
  
  colnames(s) <- stringr::str_trim(colnames(s))
  
  if ( dim(s)[1] ==6 ) { rownames(s) <- c('Min','Q1','Median','Mean','Q3','Max') 
  } else if ( dim(s)[1] ==7 ) { rownames(s) <- c('Min','Q1','Median','Mean','Q3','Max','NA') }
  
  xkabledply(s, title=title, digits = digits, pos=pos, bso=bso )
}

xkablevif = function(model, title="VIFs of the model", digits = 3, pos="left", bso="striped", wide=TRUE) { 
  #' Combining faraway::vif, xtable, and kableExtra, to easily display numeric summary of VIFs for a model. 
  #' ELo 202004 GWU DATS
  #' version 1.2
  #' @param model The lm or compatible model object.
  #' @param title Title of table. 
  #' @param digits Number of digits to display
  #' @param pos Position of table, c("left","center","right") 
  #' @param bso bootstrap_options = c("basic", "striped", "bordered", "hover", "condensed", "responsive")
  #' @param wide print table in long (FALSE) format or wide (TRUE) format
  #' @return The HTML summary table of the VIFs for a model for display, or for knitr to process into other formats 
  #' @examples
  #' xkablevif( lm(Salary~Hits+RBI, data=ISLR::Hitters), wide=T )
  
  vifs = table( names(model$coefficients)[2:length(model$coefficients)] ) # remove intercept to set column names
  vifs[] = faraway::vif(model) # set the values
  if (wide) { vifs <- t(vifs) }
  xkabledply( vifs, title=title, digits = digits, pos=pos, bso=bso )
}
```

# NY Child Care 

## EAD

### Step 1  
**Import the data, call it `childcare_orig.`**  
```{r S1}
loadPkg("readr")
childcare_orig <- data.frame(read_csv("C:\\Users\\ashwi\\Desktop\\DOHMH_Childcare_Center_Inspections.csv"))
nrow(childcare_orig)
```


### Step 2
**Remove redundant duplicated entries, drop unwanted variables, call it childcare**
```{r}
childcare <- childcare_orig[!duplicated(childcare_orig), ] %>%
             subset(select = -c(ZipCode, Building, Street, Phone, Permit.Number, Day.Care.ID, Building.Identification.Number, Actual, URL, Regulation.Summary, Health.Code.Sub.Section, Inspection.Summary.Result, Age.Range))

childcare$Program.Type = toupper(childcare$Program.Type)

nrow(childcare)
```
We can see there were quite a lot human error in this dataset (over 8000 unnecessary duplicate entry)


### Step 3
**Select only the unique rows based on Names, Call it Centers**
```{r}
Centers <- childcare[!duplicated(childcare$Center.Name), ]
nrow(Centers)
```

There are 2211 different childcare institutions in this dataset.



###Step 4
**Correlation Matrix for numeric variables**


```{r}
library(corrplot)
violation = childcare %>% subset(select = c(Violation.Rate.Percent, Total.Educational.Workers, Public.Health.Hazard.Violation.Rate, Critical.Violation.Rate, Maximum.Capacity))
corr_violation = cor(violation, use = "complete.obs")
corrplot(corr_violation, method = 'number')
```

#Step 5 - Generalized linear model for predicting if inspection will lead to a violation being cited or not

```{r}

loadPkg("dplyr")

# Remove unnecessary columns and create column called Violation.Flag to show if inspection has resulted in a violation or not

glm_childcare <- childcare %>% subset(select = -c(Date.Permitted))

glm_childcare = glm_childcare %>% mutate(Violation.Flag = ifelse(is.na(Violation.Category), 0, 1))

# Remove/replace NAs with desired values in each column

glm_childcare$Violation.Category[is.na(glm_childcare$Violation.Category)] <- 'NO VIOLATION'
glm_childcare$Violation.Rate.Percent[is.na(glm_childcare$Violation.Rate.Percent)] <- 0
glm_childcare$Average.Violation.Rate.Percent[is.na(glm_childcare$Average.Violation.Rate.Percent)] <- 0
glm_childcare$Public.Health.Hazard.Violation.Rate[is.na(glm_childcare$Public.Health.Hazard.Violation.Rate)] <- 0
glm_childcare$Average.Public.Health.Hazard.Violation.Rate[is.na(glm_childcare$Average.Public.Health.Hazard.Violation.Rate)] <- 0
glm_childcare$Violation.Rate.Percent[is.na(glm_childcare$Violation.Rate.Percent)] <- 0
glm_childcare$Average.Violation.Rate.Percent[is.na(glm_childcare$Average.Violation.Rate.Percent)] <- 0
glm_childcare$Critical.Violation.Rate[is.na(glm_childcare$Critical.Violation.Rate)] <- 0
glm_childcare$Average.Critical.Violation.Rate[is.na(glm_childcare$Average.Critical.Violation.Rate)] <- 0
glm_childcare <- glm_childcare %>% filter(!is.na(Inspection.Date))

# Check if dataset has any NA values

sapply(glm_childcare,function(x) sum(is.na(x)))

# Convert categorical variables to required datatype

glm_childcare$Program.Type <- as.factor(glm_childcare$Program.Type)
glm_childcare$Borough <- as.factor(glm_childcare$Borough)
glm_childcare$Status <- as.factor(glm_childcare$Status)
glm_childcare$Program.Type <- as.factor(glm_childcare$Program.Type)
glm_childcare$Facility.Type <- as.factor(glm_childcare$Facility.Type)
glm_childcare$Child.Care.Type <- as.factor(glm_childcare$Child.Care.Type)
glm_childcare$Violation.Category <- as.factor(glm_childcare$Violation.Category)
glm_childcare$Violation.Status <- as.factor(glm_childcare$Violation.Status)
glm_childcare$Violation.Flag <- as.factor(glm_childcare$Violation.Flag)

# Randomize test and train datasets (90:10)

glm_childcare
require(caTools)
set.seed(1000) 
sample = sample.split(glm_childcare, SplitRatio = .90)
train = subset(glm_childcare, sample == TRUE)
test  = subset(glm_childcare, sample == FALSE)

# GLM model training

model <- glm(Violation.Flag ~ Public.Health.Hazard.Violation.Rate + Critical.Violation.Rate + Borough, family=binomial, data=train)

summary(model)

# VIF check for collinearity (Remove multicollinear factors eg: Violation.Rate.Percent has a strong positive correlation with Critical.Violation.Rate.Percent)

xkablevif(model)

# Model Accuracy calculated using test dataset

fitted.results <- predict(model,newdata=subset(test,select=c(14,16,3)),type='response')
fitted.results <- ifelse(fitted.results > 0.49,1,0)

misClasificError <- mean(fitted.results != test$Violation.Flag)
print(paste('Accuracy',1-misClasificError))

```

