---
title: "NY Child Care_EAD"
author: 
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    # number_sections: true
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r basic, include=F}
# use this function to conveniently load libraries and work smoothly with knitting
# can add quietly=T option to the require() function
# the loadPkg function essentially replaced/substituted two functions install.packages() and library() in one step.
loadPkg = function(x) { if (!require(x,character.only=T, quietly =T)) { install.packages(x,dep=T,repos="http://cran.us.r-project.org"); if(!require(x,character.only=T)) stop("Package not found") } }

# unload/detact package when done using it
unloadPkg = function(pkg, character.only = FALSE) { 
  if(!character.only) { pkg <- as.character(substitute(pkg)) } 
  search_item <- paste("package", pkg,sep = ":") 
  while(search_item %in% search()) { detach(search_item, unload = TRUE, character.only = TRUE) } 
}
```

```{r setup, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
knitr::opts_chunk$set(warning = F, results = "markup", message = F)
# knitr::opts_chunk$set(warning = F, results = "hide", message = F)
# knitr::opts_chunk$set(include = F)
# knitr::opts_chunk$set(echo = TRUE)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

```{r xkablesummary}
loadPkg("xtable")
loadPkg("kableExtra")
loadPkg("stringi")

xkabledply = function(modelsmmrytable, title="Table", digits = 4, pos="left", bso="striped", wide=FALSE) { 
  #' Combining base::summary, xtable, and kableExtra, to easily display model summary. 
  #' wrapper for the base::summary function on model objects
  #' Can also use as head for better display
  #' ELo 202004 GWU DATS
  #' version 1.2
  #' @param modelsmmrytable This can be a generic table, a model object such as lm(), or the summary of a model object summary(lm()) 
  #' @param title Title of table. 
  #' @param digits Number of digits to display
  #' @param pos Position of table, c("left","center","right") 
  #' @param bso bootstrap_options = c("basic", "striped", "bordered", "hover", "condensed", "responsive")
  #' @param wide print table in long (FALSE) format or wide (TRUE) format
  #' @return HTML table for display
  #' @examples
  #' library("xtable")
  #' library("kableExtra")
  #' xkabledply( df, title="Table testing", pos="left", bso="hover" )
  #' xkabledply( ISLR::Hitters[1:5,] )
  if (wide) { modelsmmrytable <- t(modelsmmrytable) }
  modelsmmrytable %>%
    xtable() %>% 
    kable(caption = title, digits = digits) %>%
    kable_styling(bootstrap_options = bso, full_width = FALSE, position = pos)
}

xkabledplyhead = function(df, rows=5, title="Head", digits = 4, pos="left", bso="striped") { 
  xkabledply(df[1:rows, ], title, digits, pos, bso, wide=FALSE)
}

xkabledplytail = function(df, rows=5, title="Tail", digits = 4, pos="left", bso="striped") { 
  trows = nrow(df)
  xkabledply(df[ (trows-rows+1) : trows, ], title, digits, pos, bso, wide=FALSE)
}

xkablesummary = function(df, title="Table: Statistics summary.", digits = 4, pos="left", bso="striped") { 
  #' Combining base::summary, xtable, and kableExtra, to easily display numeric variable summary of dataframes. 
  #' ELo 202004 GWU DATS
  #' version 1.2
  #' @param df The dataframe.
  #' @param title Title of table. 
  #' @param digits Number of digits to display
  #' @param pos Position of table, c("left","center","right") 
  #' @param bso bootstrap_options = c("basic", "striped", "bordered", "hover", "condensed", "responsive")
  #' @return The HTML summary table for display, or for knitr to process into other formats 
  #' @examples
  #' xkablesummary( faraway::ozone )
  #' xkablesummary( ISLR::Hitters, title="Five number summary", pos="left", bso="hover"  )
  
  s = summary(df) %>%
    apply( 2, function(x) stringr::str_remove_all(x,c("Min.\\s*:\\s*","1st Qu.\\s*:\\s*","Median\\s*:\\s*","Mean\\s*:\\s*","3rd Qu.\\s*:\\s*","Max.\\s*:\\s*")) ) %>% # replace all leading words
    apply( 2, function(x) stringr::str_trim(x, "right")) # trim trailing spaces left
  
  colnames(s) <- stringr::str_trim(colnames(s))
  
  if ( dim(s)[1] ==6 ) { rownames(s) <- c('Min','Q1','Median','Mean','Q3','Max') 
  } else if ( dim(s)[1] ==7 ) { rownames(s) <- c('Min','Q1','Median','Mean','Q3','Max','NA') }
  
  xkabledply(s, title=title, digits = digits, pos=pos, bso=bso )
}

xkablevif = function(model, title="VIFs of the model", digits = 3, pos="left", bso="striped", wide=TRUE) { 
  #' Combining faraway::vif, xtable, and kableExtra, to easily display numeric summary of VIFs for a model. 
  #' ELo 202004 GWU DATS
  #' version 1.2
  #' @param model The lm or compatible model object.
  #' @param title Title of table. 
  #' @param digits Number of digits to display
  #' @param pos Position of table, c("left","center","right") 
  #' @param bso bootstrap_options = c("basic", "striped", "bordered", "hover", "condensed", "responsive")
  #' @param wide print table in long (FALSE) format or wide (TRUE) format
  #' @return The HTML summary table of the VIFs for a model for display, or for knitr to process into other formats 
  #' @examples
  #' xkablevif( lm(Salary~Hits+RBI, data=ISLR::Hitters), wide=T )
  
  vifs = table( names(model$coefficients)[2:length(model$coefficients)] ) # remove intercept to set column names
  vifs[] = faraway::vif(model) # set the values
  if (wide) { vifs <- t(vifs) }
  xkabledply( vifs, title=title, digits = digits, pos=pos, bso=bso )
}
```

```{r outlierKD2}
# Fix outliers
outlierKD2 <- function(df, var, rm=FALSE, boxplt=FALSE, histogram=TRUE, qqplt=FALSE) { 
    #' Original outlierKD functino by By Klodian Dhana,
    #' https://www.r-bloggers.com/identify-describe-plot-and-remove-the-outliers-from-the-dataset/
    #' Modified to have third argument for removing outliers instead of interactive prompt, 
    #' and after removing outlier, original df will not be changed. The function returns the a df, 
    #' which can be saved as original df name if desired.
    #' Also added QQ-plot in the output, with options to show/hide boxplot, histogram, qqplot.
    #' Check outliers, and option to remove them, save as a new dataframe. 
    #' @param df The dataframe.
    #' @param var The variable in the dataframe to be checked for outliers
    #' @param rm Boolean. Whether to remove outliers or not.
    #' @param boxplt Boolean. Whether to show the boxplot, before and after outliers removed.
    #' @param histogram Boolean. Whether to show the histogram, before and after outliers removed.
    #' @param qqplt Boolean. Whether to show the qqplot, before and after outliers removed.
    #' @return The dataframe with outliers replaced by NA if rm==TRUE, or df if nothing changed
    #' @examples
    #' outlierKD2(mydf, height, FALSE, TRUE, TRUE, TRUE)
    #' mydf = outlierKD2(mydf, height, TRUE, TRUE, TRUE, TRUE)
    #' mydfnew = outlierKD2(mydf, height, TRUE)
    dt = df # duplicate the dataframe for potential alteration
    var_name <- eval(substitute(var),eval(dt))
    na1 <- sum(is.na(var_name))
    m1 <- mean(var_name, na.rm = T)
    par(mfrow=c(2, boxplt+histogram+qqplt), oma=c(0,0,3,0))
    if (qqplt) { 
      qqnorm(var_name, main = "With outliers")
      qqline(var_name)
    }
    if (histogram) { hist(var_name, main="With outliers", xlab=NA, ylab=NA) }
    if (boxplt) { boxplot(var_name, main="With outliers") }

    outlier <- boxplot.stats(var_name)$out
    mo <- mean(outlier)
    var_name <- ifelse(var_name %in% outlier, NA, var_name)
    if (qqplt) { 
      qqnorm(var_name, main = "Without outliers")
      qqline(var_name)
    }
    if (histogram) { hist(var_name, main="Without outliers", xlab=NA, ylab=NA) }
    if (boxplt) { boxplot(var_name, main="Without outliers") }

    title("Outlier Check", outer=TRUE)
    na2 <- sum(is.na(var_name))
    cat("Outliers identified:", na2 - na1, "\n")
    cat("Propotion (%) of outliers:", round((na2 - na1) / sum(!is.na(var_name))*100, 1), "\n")
    cat("Mean of the outliers:", round(mo, 2), "\n")
    m2 <- mean(var_name, na.rm = T)
    cat("Mean without removing outliers:", round(m1, 2), "\n")
    cat("Mean if we remove outliers:", round(m2, 2), "\n")
    
    # response <- readline(prompt="Do you want to remove outliers and to replace with NA? [yes/no]: ")
    # if(response == "y" | response == "yes"){
    if(rm){
        dt[as.character(substitute(var))] <- invisible(var_name)
        #assign(as.character(as.list(match.call())$dt), dt, envir = .GlobalEnv)
        cat("Outliers successfully removed", "\n")
        return(invisible(dt))
    } else {
        cat("Nothing changed", "\n")
        return(invisible(df))
    }
}
```
# NY Child Care 

## EAD

### Step 1  
**Import the data, call it `childcare_orig.`**  
```{r S1}
loadPkg("readr")
childcare_orig <- data.frame(read_csv("DOHMH_Childcare_Center_Inspections.csv"))
nrow(childcare_orig)
```


### Step 2
**Remove redundant duplicated entries, drop unwanted variables, call it childcare**
```{r}
childcare <- childcare_orig[!duplicated(childcare_orig), ]
childcare<-subset(childcare,select = -c(ZipCode, Building, Street, Phone, Permit.Number, Day.Care.ID, Building.Identification.Number, Actual, URL, Regulation.Summary, Health.Code.Sub.Section, Inspection.Summary.Result))

childcare$Program.Type <- toupper(childcare$Program.Type)

nrow(childcare)
```
We can see there were quite a lot human error in this dataset (over 8000 unnecessary duplicate entry)


### Step 3
**Select only the unique rows based on Names, Call it Centers**

```{r}
Centers <- childcare[!duplicated(childcare$Center.Name), ]
nrow(Centers)
```

There are 2211 different childcare institutions in this dataset.



###Step 4
**Discover the distribution of Borough and Program Type across different Institutions
```{r S4_1}
loadPkg("ggplot2")
bar_Bor <- ggplot(Centers, aes(x = Borough)) +
           geom_bar() +
           geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5)
bar_Bor
```

```{r}
loadPkg("dplyr")
Bor <- group_by(Centers, Borough) %>%
       count() %>%
       ungroup()%>% 
       arrange(desc(Borough)) %>%
       mutate(percentage = round(n/sum(n), 4) * 100,
       lab.pos = cumsum(percentage)-.5*percentage,
       ymax = cumsum(percentage),
       ymin = c(0, head(ymax, n = -1)))

donut_Bor <- ggplot(Bor, aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3,fill = Borough)) +
             geom_rect() +
             geom_text(x = 2, aes(y = lab.pos, label = percentage, color = Borough), size = 5) +
             scale_fill_brewer(palette = 4) +
             scale_color_brewer(palette = 4) +
             coord_polar(theta = "y") +
             xlim(c(-1, 4)) +
             theme_void() 
donut_Bor
```



```{r S4_2}
bar_Type <- ggplot(Centers, aes(x = Program.Type)) +
            geom_bar() +
            geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5)
bar_Type
```

```{r}
Type <- group_by(Centers, Program.Type) %>%
        count() %>%
        ungroup()%>% 
        arrange(desc(Program.Type)) %>%
        mutate(percentage = round(n/sum(n), 4) * 100,
        lab.pos = cumsum(percentage)-.5*percentage,
        ymax = cumsum(percentage),
        ymin = c(0, head(ymax, n = -1)))

donut_Type <- ggplot(Type, aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3, fill = Program.Type)) +
              geom_rect() +
              geom_text(x = 2, aes(y = lab.pos, label = percentage, color = Program.Type), size = 5) +
              scale_fill_brewer(palette = 3) +
              scale_color_brewer(palette = 3) +
              coord_polar(theta = "y") +
              xlim(c(-1, 4)) +
              theme_void() 
donut_Type
```

Preschool Camp and School Age Camp have only 1 observation each. Consider to drop them later.


```{r}
bar_VioCat <- ggplot(childcare, aes(x = Violation.Category)) +
            geom_bar() +
            geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5)
bar_VioCat
```

```{r}
VioCat <- group_by(childcare, Violation.Category) %>%
          count() %>%
          ungroup()%>% 
          arrange(Violation.Category) %>%
          mutate(percentage = round(n/sum(n), 4) * 100,
          lab.pos = cumsum(percentage)-.5*percentage,
          ymax = cumsum(percentage),
          ymin = c(0, head(ymax, n = -1)))

donut_VioCat <- ggplot(VioCat, aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3, fill = Violation.Category)) +
                geom_rect() +
                geom_text(x = 2, aes(y = lab.pos, label = percentage, color = Violation.Category), size = 5) +
                scale_fill_brewer(palette = 4) +
                scale_color_brewer(palette = 4) +
                coord_polar(theta = "y") +
                xlim(c(-1, 4)) +
                theme_void() 
donut_VioCat
```

The NA in Violation Category means no violation was detected during that inspection. Only around 44% of the inspections where violation-free.




**Histogram for Educational workers**
```{r}
hist_Edu_Bor <- ggplot(Centers, aes(x = Total.Educational.Workers, 
                                    color = Borough, 
                                    fill = Borough)) +
                geom_histogram(binwidth = 1,
                               alpha = 0.3, 
                               position = "identity")+ 
                theme(plot.title = element_text(hjust = 0.5)) + 
                labs(title = "Histogram for No. of Educational Workers (Colored by Borough)", 
                     x = "No. of Educational Workers",
                     y = "Count")
hist_Edu_Bor
```

```{r}
hist_Edu_Typ <- ggplot(Centers, aes(x = Total.Educational.Workers, 
                                    color = Program.Type, 
                                    fill = Program.Type)) +
                geom_histogram(binwidth = 1,
                               alpha = 0.3, 
                               position = "identity")+ 
                theme(plot.title = element_text(hjust = 0.5)) + 
                labs(title = "Histogram for No. of Educational Workers (Colored by Program Type)", 
                     x = "No. of Educational Workers",
                     y = "Count")
hist_Edu_Typ
```

***Above plot might not be useful, but I put it here anyway***


**Boxplot for Educational workers**
```{r}
box_Edu_Bor <- ggplot(Centers, aes(y = Total.Educational.Workers,
                                   x = Borough,
                                   fill = Borough, 
                                   color = Borough)) + 
               geom_boxplot(alpha = 0.3) +
               theme(plot.title = element_text(hjust = 0.5)) +
               ggtitle("Boxplot for No. of Educational workers (Colored by Borough)") +
               scale_y_continuous(name = "No. of Educational workers") + 
               scale_x_discrete(name = "Borough")
box_Edu_Bor
```



```{r}
box_Edu_Typ <- ggplot(Centers, aes(y = Total.Educational.Workers,
                                   x = Program.Type,
                                   fill = Program.Type, 
                                   color = Program.Type)) + 
               geom_boxplot(alpha = 0.3) +
               theme(plot.title = element_text(hjust = 0.5)) +
               ggtitle("Boxplot for No. of Educational workers (Colored by Program Type)") +
               scale_y_continuous(name = "No. of Educational workers") + 
               scale_x_discrete(name = "Program Type")
box_Edu_Typ
```

***Above plot might not be useful, but I put it here anyway***

```{r}
median(Centers$Total.Educational.Workers)
quantile(Centers$Total.Educational.Workers, 0.8)
```



**Histogram for Maximum Capacity**
```{r}
hist_Cap_Bor <- ggplot(Centers, aes(x = Maximum.Capacity, 
                                    color = Borough, 
                                    fill = Borough)) +
                geom_histogram(binwidth = 5,
                               alpha = 0.3, 
                               position = "identity")+ 
                theme(plot.title = element_text(hjust = 0.5)) + 
                labs(title = "Histogram for Maximum Capacity (Colored by Borough)", 
                     x = "Maximum Capacity",
                     y = "Count")
hist_Cap_Bor
```


```{r}
box_Cap_Bor <- ggplot(Centers, aes(y = Maximum.Capacity,
                                   x = Borough,
                                   fill = Borough, 
                                   color = Borough)) + 
               geom_boxplot(alpha = 0.3) +
               theme(plot.title = element_text(hjust = 0.5)) +
               ggtitle("Boxplot for Maximum Capacity (Colored by Borough)") +
               scale_y_continuous(name = "Maximum Capacity") + 
               scale_x_discrete(name = "Borough")
box_Cap_Bor
```

```{r}
median(Centers$Maximum.Capacity)
quantile(Centers$Maximum.Capacity, 0.8)
```

From the histogram and the box plot, the big picture for childcare industry is portrayed: The largest bulk consists of small care units with 15 educational workers or less, tending less then 100 children maximum. These institutions serve their local community. But like all education facilities, different needs are always to be tailored and catered. The presence of larger institution with more workers, taking care of more children persists, making the entire distribution long-tailed.

Borough-wise, we can see Staten Island has fewest childcare institutions, while Brooklyn have the most. This could be explained by the difference in population across different Boroughs. Manhattan, which is the highest GDP per capita Borough in NY, does not have a large number of childcare center. This observation could attribute to demographic and income structure: Manhattan is the Financial and Commercial center, but is not a residential center, as people tends to live in other parts of the city due to land price. For people who resides in this Borough who are more likely to have higher income to afford the cost of living, they have more alternatives for their children including domestic nurses and tutoring, boarding schools etc. Resulting in less demand for childcare center in this area. This could be contrasted in Brooklyn, where the Big Apple's working class work and live. Here people are more likely to require a childcare center since they have less alternative for their children after school.


```{r}
normworkers<-shapiro.test(Centers$Total.Educational.Workers)
normworkers
```


```{r}
box_Edu_Bor <- ggplot(Centers, aes(y = Total.Educational.Workers,
                                   x = Borough,
                                   fill = Borough, 
                                   color = Borough)) + 
               geom_boxplot(alpha = 0.3) +
               theme(plot.title = element_text(hjust = 0.5)) +
               ggtitle("Boxplot for No. of Educational workers (Colored by Borough)") +
               scale_y_continuous(name = "No. of Educational workers") + 
               scale_x_discrete(name = "Borough")
```


```{r}
boxplot(Centers$Total.Educational.Workers)
outliers<-boxplot(Centers$Total.Educational.Workers)$out
x<-Centers
x<-x[-which(x$Total.Educational.Workers %in% outliers),]
boxplot(x$Total.Educational.Workers)
```

```{r}
test<-shapiro.test(x$Total.Educational.Workers)
test
```
```{r}
x$Borough<-as.factor(x$Borough)
```


```{r}
kruskal.test(g=x$Borough,x=x$Total.Educational.Workers)
install.packages("FSA")
library(FSA)
```


```{r}
d <- dunnTest(Total.Educational.Workers~ Borough, data = x)
d
```
```{r}
child2<-childcare[c("Inspection.Date","Violation.Rate.Percent","Borough")]
child2 <- na.omit(child2)
```

```{r}
child2$Inspection.Date<-substring(child2$Inspection.Date,nchar(as.character(child2$Inspection.Date))-3)

```

```{r}
loadPkg("dplyr")
d1<-child2 %>% group_by(Borough,Inspection.Date)%>% mutate(Violation.Rate.Percent=mean(Violation.Rate.Percent))
unique(d1)
```

```{r}
d1$Inspection.Date<-as.numeric(d1$Inspection.Date)
d1$Violation.Rate.Percent<-as.numeric(d1$Violation.Rate.Percent)
d1brokk<-subset(d1,Borough=="BROOKLYN")
d1bqn<-subset(d1,Borough=="QUEENS")
d1st<-subset(d1,Borough=="STATEN ISLAND")
d1mn<-subset(d1,Borough=="MANHATTAN")
d1brons<-subset(d1,Borough=="BRONX")
c1<-cor.test(d1brokk$Inspection.Date,d1brokk$Violation.Rate.Percent, na.action = na.omit)
c1

```
```{r}
c2<-cor.test(d1bqn$Inspection.Date,d1bqn$Violation.Rate.Percent, na.action = na.omit)
c2
```

```{r}
c3<-cor.test(d1st$Inspection.Date,d1st$Violation.Rate.Percent, na.action = na.omit)
c3
```

```{r}
c4<-cor.test(d1mn$Inspection.Date,d1mn$Violation.Rate.Percent, na.action = na.omit)
c4
```

```{r}
c5<-cor.test(d1brons$Inspection.Date,d1brons$Violation.Rate.Percent, na.action = na.omit)
c5
```
```{r}
install.packages("ggthemes") # Install 
library(ggthemes)
```


```{r}
loadPkg("ggplot2")

ggplot(d1,aes(x=Inspection.Date,y=Violation.Rate.Percent,group=Borough,color=Borough))+geom_point(size=2,color='red')+geom_line(size=1)+ggtitle("Violation Rate percent vs. Inspection date across Borough")+theme_stata() + scale_color_stata() 
```

Which borough has improved over the last four years?
Here, the improvement means the decrease in violation rate as the inspection year increases. The above cor.test() results shows that Queens and Manhattan has good correlation, which says that the violation rate is decreasing over the years.The visualization above proves the test results.
